Automated Synthesis of State-Space Digital Systems

### 2016-12-19 cav2017/cav2017.tex
Early draft on The state-space model

\usetikzlibrary{positioning, arrows, automata, shapes}

\newcommand{\todo}[1]{{\color{red} TODO: #1}}
\newcommand{\commentib}[1]{{\color{blue} [IB: #1]}}

@@

In this paper we formalise the notion of solution generalisation for CEGIS. 
This technique allows the inductive generalisation engine to look for 
candidate solutions in a reduced solution space.

@@

%-------------------------------
\section{Discretising Continuous Space Dynamics}
%-------------------------------

Discretization of an continuous dynamical system turns the ODE into a difference equation, assuming zero-order hold
for the input $\vec{u}$ and continuous integration for the noise $\vec{w}$, to
\begin{align}
\label{eq:discretization}
\vec{x}_{k+1} &= \vec{A}_d\vec{x}_k+\vec{B}_d\vec{u}_k + \vec{w}_k,\\
y_k &= \vec{C}_d \vec{x}_ k + \vec{D}_d \vec{u}_ k. 
\end{align}
with covariance for $\vec{w}_k \sim \mathcal{N}(0,Q_d)$,
where
\begin{align}
\label{eq:discretize}
\vec{A}_d &= e^{\vec{A} T_s} = \mathcal{L}^{-1} { ( s \vec{I} - \vec{A} )^{-1} }_{t = T_s},\\
\vec{B}_d &= \int_{0}^{T_s} e^{\vec{A} t} dt\ \vec{B} = \vec{A}^{-1} ( \vec{A}_d - \vec{I} ) \vec{B},\\
\vec{C}_d &= \vec{C},\\
\vec{D}_d &= \vec{D},\\
Q_d &= \int_{0}^{T_s} e^{\vec{A} t} Q e^{\vec{A}^T t} dt.
\end{align}
and $T_s$ is the sample time. Then
\begin{align*}
x(kT)=x_k \wedge y(kT) = y_k, \forall k.
\end{align*}
Let $g_d(k)=\mathcal{G}(t,g(t),T)$ be a function that performs the discretization described above, where $g(t)$
represents the continuous dynamics and $g_d(k)$ the corresponding discrete dynamics. 
Let $G(s)=\mathcal{L}(g(t))$ and $G_d(z)=\mathcal{Z}(g_d(k))$ be the corresponding Laplace and Z-transforms
of $g(t)$ and $g_d(k)$. Given this relation, we have 
$$G_d(z)=G(z)|_{z=e^{sT}} : g_d(k)=\mathcal{G}(t,g(t),T) \wedge T < \frac{1}{.5f_s},$$
where $.5f_s$ is the Nyquist frequency of $g(t)$. This last restriction is introduced to avoid the effects of aliasing
which could cause ``phantom poles'' to appear otherwise. 
The eigenvalues of $\vec{A}_d$ corresponding to the poles of $G_d(z)$, and those of $\vec{A}$ corresponding to the poles of $G(s)$ are similarly related.
$\hat{\lambda}_i=e^{-\lambda_iT} : \hat{\lambda}_i \in \sigma(\vec{A}_d) \wedge \lambda_i \in \sigma(\vec{A})$
where $\sigma(\cdot)$ is the spectrum of a matrix.
The following remark is worth mentioning regarding the above.
\begin{remark}
The witnessed maximum amplitude of the discrete signals $x_k,y_k$ may be smaller to that of $x(t),y(t)$ due to synchronism at fraction-frequency sampling.
This means that reasoning about the state-space must consider the effects of this possibility as well of the case of maximal inputs from the continuous specification.
\end{remark}
\textcolor{red}{LC: Dario, where is the section you mention?... Since \eqref{eq:discretization} is a bisimulation of \eqref{eq:dynamical}, we may use the semantics in section \ref{sec:model_semantics} to model continuous dynamical systems.}

%-------------------------------
\section{LTI representations of ARMA models}
%-------------------------------

Let us have a discrete time SISO system with an $n^{th},m{^th} : n\geq m$ order armax model
$$y[k]=\sum_{i=1}^n a_iy[k-i]+\sum_{i=0}^m b_iu[k-i],$$
where $y$ is the output of the system and $u$ its input.
The equivalent LTI dynamics of such a model are:
\begin{align}
\vec{x}_k=&\vec{A}\vec{x}_{k-1}+\vec{B}u_k : \vec{x}_k=[y_{k-n}\ \hdots y_{k-1}\ y_{k}]^T\\
y_k=&\vec{C}\vec{x}_k + b_0u_k\nonumber\\
\vec{A}=&\left[
\begin{array}{ccccc}
0&1&0&\cdots&0\\
0&0&1&\cdots&0\\
\vdots&\vdots&\vdots&\ddots&\vdots\\
0&0&0&\cdots&1\\
-a_n&-a_{n-1}&-a_{n-2}&\cdots&-a_1
\end{array}\right],
\vec{B}=\left[
\begin{array}{c}
0\\0\\ \vdots\\ 0\\ 1
\end{array}\right]\nonumber\\
\vec{C}=&[b_n-a_nb_0\ b_{n-1}-a_{n-1}b_0 \cdots b_1-a_1b_0] : b_{i \in (m\ n]}=0\nonumber
\end{align}

@@

>>>>

\usetikzlibrary{positioning, automata, shapes.arrows, calc, shapes, arrows}

\newcommand{\todo}[1]{{\color{red} TODO: #1}}
\newcommand{\commentib}[1]{{\color{blue} [IB: #1]}}
\newcommand{\mat}[1]{\boldsymbol{#1}}

@@

Linear Time Invariant Control systems represnt a sizeable portion of modern
embedded software, performing highly non-trivial control
tasks, with significant impact in numerous application areas such as
environmental control, navigation and robotics~\cite{astrom1997computer,Franklin15}.

Correct and sound control is non-trivial, however. The problem is exacerbated by
artefacts specific to digital control, such as the effects of
finite-precision arithmetic, time discretization, and quantization noise
introduced by A/D and D/A conversion.  Thus, the programming is a key
barrier to broad adoption of digital control, and requires considerable
expertise.

Whilst there is a vast literature in validation of control systems, the subject
of control synthesis has been somewhat restricted in the properties it explores.
Traditionally, this focuses mainly on the subject of stability, working typically on
continuous domains \cite{sadabadi2016static}, or using switched models~\cite{DBLP:conf/emsoft/RavanbakhshS16}.
Reachability tools are also used when discretizing the state-space for both verification and parametric synthesis\cite{cimatti2013parameter}.

In this paper, we propose the synthesis of control algorithms for cyberphisical
systems taking into consideration the continuous domain of the plant alongside
the discrete domain of the controller and the hybrid elements that communicate
them.

To this end, we formalise the notion of solution generalisation for CEGIS and 
apply CEGAR on simplified abstractions of the system in order to achieve scalability. 
The former technique allows the inductive generalisation engine to look for 
candidate solutions in a reduced solution space. 
The latter case is a well know technique for abstract interpretation that allows
simplification of the procedures over a sound abstraction of the verification model.

%-------------------------------
\section{Preliminaries}
%-------------------------------
In this work we will focus on Linear Dynamical Cyberphisical Systems with
continuous plants described by ODEs in the form
\begin{align}
\label{eq:ode}
\dot{\vec{x}}(t)&=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)\\
\vec{y}(t)&=\mat{C}\vec{x}+\mat{D}\vec{u}\nonumber
\end{align}
which are soundly discretized (as shown in the appendix) into 
\begin{align}
\label{eq:plant}
\vec{x}_{k+1}&=\mat{A}_d\vec{x}_k+\mat{B}_d\vec{u}_k\\
\vec{y}_k&=\mat{C}_d\vec{x}_k+\mat{D}_d\vec{u}_k\nonumber
\end{align}
and Digital Controllers in the form
\begin{equation}
\vec{u}_k=\mat{K}_d\vec{y}_k
\end{equation}
The output of the plant is being sampled by an ADC, and since the full state of the
plant cannot be witnessed directly, the controller software includes an observer that 
estimates the state of the plant (see appendix for details on how such an observer
is constructed).

We further presume that there are small tolerances in the system described in the
following form
\begin{align*}
\mat{A} \in \mathcal{G}_A=\mat{S}\mat{J}\mat{S}^{-1} : \mat{J} \in \mathcal{G}_J
\mat{B} \in \mathcal{G}_B
\end{align*}

Finally, we also require a safety specification $\Psi(\vec{u},\vec{y})$ which is
given as a set of linear constraints over the inputs and outputs.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Counterexample Guided Abstractly Refined Inductive Synthesis for LTI Cyberphysical Systems}\label{sec:CEGARIS}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Verification of a Conrollable Observed System}\label{sec:cof_verification}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

Let us recall the formulas used for SISO systems in sections \ref{sec:reachable} and \ref{sec:observable}.
A controllable and observable system will have the following dynamics:
\begin{align}
\label{eq:observer_LTI_cf}
\left [\begin{array}{c}\vec{x}\\ \hat{\vec{x}}_e \end{array}\right]_{k+1}
=&\left [\begin{array}{cc}\mat{A}_{tp}&\mat{B}_{to}\\ \mat{0}&\mat{A}_{to}\end{array}\right]
\left [\begin{array}{c}\vec{x}\\ \hat{\vec{x}}_e \end{array}\right]_k
+\mat{B}_{tp} \vec{r}_k
+\mat{B}_{tn}\left [\begin{array}{c}\vec{\nu}_2\\ \vec{\nu}_1\end{array}\right]\\
\vec{y}_k=&\mat{C}_{t}\vec{x}_k + \mat{D}_{d}\vec{u}_k\nonumber
\end{align}
\begin{align*}
\mat{A}_{tp}=&\mat{T}\mat{A}_d\mat{T}^{-1}-\mat{T}\mat{B}_d\mat{K}_d\mat{T}^{-1}&
\mat{A}_{to}=&\mat{T}\mat{A}_d\mat{T}^{-1}-\mat{L}_d\mat{C}_d\mat{T}&
\mat{A}_t=&\left [\begin{array}{cc}\mat{A}_{tp}&\mat{B}_{to}\\ \mat{0}&\mat{A}_{to}\end{array}\right]\\
\mat{B}_{to}=&\mat{T}\mat{B}_d\mat{K}_d\mat{T}^{-1}&
\mat{B}_{tp}=&\left[\begin{array}{c}\mat{T}\mat{B}_d\mat{k}_r\\\mat{0}\end{array}\right]&
\mat{B}_{tn}=&\left[\begin{array}{cc}\mat{T}\mat{B}_d\mat{k}_r&\mat{0}\\\mat{0}&-\mat{L}_d\end{array}\right]\\
\mat{T}=&\mat{R}_{cf}\mat{R}^{-1}&
\mat{L}=&\mat{W}_o^{-1}\mat{W}_{of} \left[ \begin{array}{c}p_1-a_1\\p_2-a_2\\p_n-a_n\end{array}\right]&
\mat{C}_t=&\mat{C}_d\mat{T}
\end{align*}

\begin{align*}
\mat{A}_{tp}=&\left[
\begin{array}{ccccc}
0&1&0&\cdots&0\\
0&0&1&\cdots&0\\
\vdots&\vdots&\vdots&\ddots&\vdots\\
0&0&0&\cdots&1\\
k_n-a_n&k_{n-1}-a_{n-1}&k_{n-2}-a_{n-2}&\cdots&k_1-a_1
\end{array}\right]
\label{eq:cf_SISO_2}
\end{align*}

Careful examination of these matrices reveals that because the $\mat{A}_t$ is upper block diagonal, the spectrum of $\mat{A}_t$ is equal to the union of the spectrum of $\mat{A}_{tp}$ and $\mat{A}_{to}$ which can be calculated separately. Additionally, these correspond to the roots of the polynomials
\begin{align*}
\left(z^n+\sum_{i=1}^p{(a_i-k_i)z^{n-i}}\right) \left(z^n+\sum_{i=1}^p{(p_i-a_i)z^{n-i}}\right)
\end{align*}

For a stabilised closed loop, we need these roots to remain within the unit circle (left hand side of the plane for continuous time).
Whilst this is not a sufficient condition for verifying safety properties, it is a necessary one, therefore we will review this verification process.
The direct approach would be to calculate the roots of the polynomials (\emph{ie} the eigenvalues of the matrix), but this ends up being to costly, particularly in the case of synthesis, so we use an indirect method.

Jury's criteria is a mathematical method used to verify whether the roots of the polynomial are within the unit circle by exploring solely the coefficients of the polynomial. In order to achieve this, we first create a matrix $\mat{M} \in \mathbb{R}^{(n+1) \times (n+1)}$ with zero-indexed coefficients $m_{i,j} : i,j \in [0,n],k=n-(i-1)$ in the following form:

\begin{equation}
\left[\begin{array}{ccccc}
c_0&c_1&\hdots&c_{n-1}&c_n\\
m_{\scriptscriptstyle 0,0}-m_{\scriptscriptstyle 0,n}^*\frac{m_{0,n}^*}{m_{0,0}}&m_{\scriptscriptstyle 0,1}-m_{\scriptscriptstyle 0,n-1}^*\frac{m_{0,n}^*}{m_{0,0}}&\hdots&m_{\scriptscriptstyle 0,n-1}-m_{\scriptscriptstyle 0,1}^*\frac{m_{0,n}^*}{m_{0,0}}&0\\
\vdots&\vdots&\vdots&\vdots&\vdots\\
m_{\scriptscriptstyle i-1,0}-m_{\scriptscriptstyle i-1,k-j}^*\frac{m_{i-1,k}^*}{m_{i-1,0}}&m_{\scriptscriptstyle i-1,1}-m_{\scriptscriptstyle i-1,k-j}^*\frac{m_{i-1,k}^*}{m_{i-1,0}}&\hdots&0&0\\
m_{\scriptscriptstyle n-1,0}-m_{\scriptscriptstyle n-1,1-j}^*\frac{m_{n-1,1}^*}{m_{n-1,0}}&0&\hdots&0&0\\
\end{array}\right]
\end{equation}

The first row corresponds to the coefficients of the original polynomial. Notice also that every row has a polynomial one degree lower than the previous. The sufficient and necessary conditions for the polynomial in row 0 to have all of its roots inside the unit circle are~\cite{astrom1997computer}:
\begin{align}
|c_n|<c_0&&
\sum_{i=0}^n c_i > 0&&
(-1)^n\sum_{i=0}^n c_i (-1)^{n-i}> 0&&
m_{i,0}>0
\end{align}

Notice that in both our polynomials $c_0=1$ and we only use reals.
Now that we have established convergence, we may examine the safety specification of the state space. In the case of SISO systems, this is given by a range of the input and a range of the output which are considered safe ($\vec{u}_k \in [\underline{u}\ \ \overline{u}] \wedge \vec{y}_k \in [\underline{y}\ \ \overline{y}]$). Whilst there are other properties we may include in the verification, we will solely focus on these for simplicity of the explanation.
We will assume that we do not want saturation in our system, hence the control loop has only one mode. We have also identified the quantization noise as a bounded non-deterministic input.
First we must translate the safety specification into the statespace. Replacing on the second part of equation \eqref{eq:observer_LTI_cf} we obtain
\begin{equation}
\Psi := \underline{y}-\overline{\mat{B}\vec{u}}<\mat{C}_t\vec{x}_k<\overline{y}-\underline{\mat{B}\vec{u}}
\label{eq:of_bounds}
\end{equation}
Note that we use the bounds of the input after the transformation by $\mat{B}$ since the matrix may reverse the bounds. Equation \eqref{eq:of_bounds} can be solved using linear programming as it describes a polyhedral region in $\mathbb{R}^n$.
The second step is evaluating the progression of the system over time in order to calculate $\vec{x}_k$ to validte against this specification. For this step we require an initial set from which the system progresses. Obviously, this set $\vec{x}_0$ must satisfy the specification, otherwise the system will be unsafe to begin with.
We will accept a specification in the form $\mat{E}\mat{T}\vec{x}_0<\mat{f}$. The presence of $\mat{T}$ is because this will typically be given in the original state-space.
Accelerating \eqref{eq:observer_LTI_cf}, we obtain:
\begin{align}
\label{eq:acc_observer_LTI_cf}
\left [\begin{array}{c}\vec{x}\\ \hat{\vec{x}}_e \end{array}\right]_k
=&\mat{A}_t^k
\left [\begin{array}{c}\vec{x}\\ \hat{\vec{x}}_e \end{array}\right]_0
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tp} \vec{r}_i
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tn}\left [\begin{array}{c}\vec{\nu}_2\\ \vec{\nu}_1\end{array}\right] :
 \left [\begin{array}{cc}\mat{E}\mat{T}&\mat{0}\end{array}\right] \left [\begin{array}{c}\vec{x}\\ \hat{\vec{x}}_e \end{array}\right]_0<\mat{f}
\end{align}
which would still require us to verify the system for every $k$ up to infinity. Instead we will use abstract acceleration~\cite{cattaruzza2015unbounded}, which translates the formula into
\begin{align}
\label{eq:aa_observer_LTI_cf}
\hat{X}^\#
=&\mathcal{A}_t X_0+\mathcal{B}_{tp} R + \mathcal{B}_{tn} N\\
X_0 =&\left \{ \left[\begin{array}{c}\vec{x}\\ \hat{\vec{x}}_e \end{array}\right] :
 \left [\begin{array}{cc}\mat{E}\mat{T}&\mat{0}\end{array}\right] \left [\begin{array}{c}\vec{x}\\ \hat{\vec{x}}_e \end{array}\right]_0<\mat{f} \right\}\nonumber\\
 R =&\{r : \mat{E}_r < \vec{f}_r \}\nonumber\\
 N=&\left \{ \left[\begin{array}{c}\vec{\nu}_2\\ \vec{\nu}_1\end{array}\right] : \nu_1 \in \left[-\frac{q1}{2}\ \ \frac{q1}{2}\right] \wedge \nu_2 \in \left[-\frac{q2}{2}\ \ \frac{q2}{2}\right]  \right \}\nonumber
\end{align}
where $\mathcal{A}_t=\bigcup_{k=1}^\infty \mat{A}^k, \mathcal{B}_{tp}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tp}, \mathcal{B}_{tn}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tn}$ are abstract matrices for the system in \eqref{eq:observer_LTI_cf}. It is worth noting that $r$ is a non-deterministic parametric input (\emph{ie} $\forall k>0, r_{k+1} =r_k$) whereas the set N is composed of non-deterministic variable inputs.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Numeric Representations and Soundness} \label{sec:numeric_rep}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

Let $\mathbb{R} \langle I,F \rangle$ be a fixed point domain with $I$ bits representing the integer part and $F$ bits
representing the decimal part. We will extend this notation to $\mathbb{R} \langle I,F,E \rangle$ to add a number of
leading or trailing bits when the domain requires prescaling (ie the maximum is smaller than $0.5$ or the minimum
is greater than $2$)
Let also $\mathbb{R}_* \langle M,E \rangle$ be a floating point domain with $M$ bits
mantissa and $E$ bits exponent. This domain corresponds to 
$$\mathbb{R}_* \langle M,E \rangle = \bigcup_i \mathbb{R} \langle 1,M-1,E_i \rangle : 2^{1-E} \leq E_i  \leq 2^{E-1}$$
This means that the distance between any two numbers in the domain depends on the
location of those numbers.
Since we've shown that a floating point domain is a union of a number of fixed point domains, we will continue our
reasoning on fixed point only. Our points will apply to the floating point domain by extension.
The following approximation errors will arise:
\begin{enumerate}
\item Truncation\\
Let $\mathcal{F}\langle I,F \rangle(\cdot) : \mathbb{R} \rightarrow \mathbb{R}\langle I,F \rangle, \tilde x=\mathcal{F}\langle I,F \rangle(x) = x-\delta : \delta=x\ \%_F\  c\tilde x$, where $\%_F$ is the modulus operation performed on the last bit of the representation ($2^{-F}$).\\
The number $\delta$ corresponds to the truncation error of the representation and it will propagate across operations.
\item Rounding\\
Every time an operation is performed on a number in the $\mathbb{R} \langle I,F,E \rangle$ domain, an error may be
introduced due to truncation of the result.
Let us define $c\langle i,f,e \rangle \in \mathbb{R}\langle I,F,E\rangle$, where $i \in [-2^{I-1}\ 2^{I-1}],f \in [0 1),e \in [-2^{E-1}\ 2^{E-1}]$ are its assigned values. The smallest number that can be represented for any given exponent is $c_{min(e)}=c\langle 0,2^{-F},e\rangle$.
The following errors appear in simple operations:
\begin{enumerate}
\item Addition/Substraction: $c\langle \cdot,e_1\rangle \pm c\langle \cdot,e_2\rangle=c\langle \cdot,max(e_1,e_2)\rangle + \delta_3 : |\delta_3| \leq (e_1==e_2) ? 0 : c_{min(max(e_2,e_1))}$ 
\item Multiplication: $c\langle \cdot,e_1\rangle * c\langle \cdot,e_2\rangle=c\langle \cdot,e_1+e_2\rangle + \delta_3 : |\delta_3| \leq c_{min(e_2+e_1)}$
\end{enumerate}
The bounds obtained for division are left to the reader as it will not be applied in this paper.

At this point it is worth mentioning that not all systems truncate equally. In the definition above $\delta$ may be positive
or negative, and even this decision can be made dependent on the sign of $x$. Common cases in computer systems are:
round downwards (\emph{ie} to $-\infty$), round upwards (\emph{ie} to $+\infty$), and round to zero.
Rounding errors are cumulative which means the overall error will statistically increase with every operation, thus algorithms
performing fewer operations can be more precise in this respect than iterative ones. 
\item Overflow\\
Another source of error introduced by Finite Word Length representations is that of overflow. The finite word length set a
maximum representable number in the domain ($\pm 2^{I-1}+1-2^{-F}$), which means that any number outside this range
cannot be accurately represented by the domain (\emph{eg} $\delta=x-2^{I-1}+1-2^{-F}$). In the case of floating point, a special
case for infinity has been introduced, which addresses some of this issues, but once the error has gone to infinity, the numbers
become meaningless. It is the responsibility of software designers and tools to account for overflows and either supply saturation
functions to prevent them or raise alarms in their presence. In the case of verification, the latter is seen as a safety problem.
\item Null/Zeno behaviour\\
Whilst this work does not explore the effect of Zeno behaviours, it is worth noting that this effect may also be caused by rounding
errors in floating point programs. The reason for this is that if an operation contains two numbers which belong to disjoint sub-domains
(\emph{ie} neither number can be represented in the domain of its counterpart), the resulting operation may become a no-op/skip,
subsequently destabilizing the program.
\end{enumerate}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Fixed Word Length Effects on Matrix Multiplications}\label{sec:cof_fwl}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Thus far we have presented a verification procedure for a digitalized plant in a continuous space (\emph{ie} the
range of the quatizers is still $\mathbb{R}$. The next step we must reaize is that the controller and the observer
exist in a digital program, hence their domain is $\mathbb{R}\langle I,F,E\rangle$. This has two separate effects
on the model. If we recall equation \eqref{eq:to_of}, the matrices of the observer and the controller  are not necessarily those of the plant.
In fact we have $\mat{A},\mat{B},\mat{C} \in \mathbb{R}^*$, $\mat{A}_o,\mat{B}_o,\mat{C}_o, \mat{K}_d, \mat{L}_d \in \mathbb{R}\langle I,F,E\rangle^*$
Where $*$ represents the appropriate dimension for each matrix.
Unfortunately this means that although combined matrices such as $\mat{A}_t$ are in the reals, they suffer from
the fixed word restrictions in their composing matrices.
The second effect corresponds to the rounding errors in the calculation of $\hat{x}_{k+1}, u_k$. These will introduce a new source of quantiztion noise which we may denote $\vec{\nu}_3 \in [-q_3\ \ q_3]$.
The estimation of $\hat{\vec{x}}$ corresponds to the sum of three matrix-vector multiplications (we assume $\mat{C}_o$ in Objective form). Each of these has a coefficient-wise error bound $c_{min(max(e_{j1}+e_{j2}))}$. Hence the overall $\overline{\delta}_{\hat{x}}=3nc_{min(max(e_{j1}+e_{j2}))}$. The error in $\vec{u}$ corresponds of the multiplication of $\mat{K}_d\hat{\vec{x}}$, hence $q_3=\overline{\delta}_u=nc_{min(max(e_{k1}+e_{k2}))}+\sum_{k=1}^nc_k\overline{\delta}_{\hat{x}})=3n|\mat{K}|_1c_{min(max(e_{j1}+e_{j2}))}+nc_{min(max(e_{k1}+e_{k2}))}$

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Synthesisig the dynamics}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Now that we have explained the verification process, we may proceed to describe the synthesis procedure, which is illustrated in figure \ref{fig:CEGARIS}.
\begin{enumerate}
\item We start by preprocessing the dynamics. This can be computationally intensive and we don't want it to be part of the CEGIS loop.
\begin{enumerate}
\item Calculate $\mat{A}_o,\mat{B}_o,\mat{C}_o$ by rounding down the coefficients $*_{oij}=\mathcal{F}\langle I,F,E\rangle (*_{ij}) : *=\{a,b,c\}$
\item Calculate $\mat{T}$ and $\mat{W}_o^{-1}\mat{W}_{of}$ using equations \eqref{eq:rncf}\eqref{eq:rcf}\eqref{eq:to_cf}\eqref{eq:wnof}\eqref{eq:wof}
\item Extract the Plant's characteristic polynomial coefficients from the last row of $P_a=\mat{T}\mat{A}\mat{T}^-1$.
\item Calculate the noise set $N$ from the quantizer resolutions and estimated round-off errors ($\frac{q_1}{2},\frac{q_2}{2},q_3$)
\end{enumerate}
\item Next we synthesise a candidate controller and observer. Let $\mathcal{J}\langle I,F,E \rangle (P,\mat{T})$ be a program describing Jury's method.
\begin{enumerate}
\item Select a controller $\tilde{\mat{K}} \in \mathbb{R}\langle I,F,E\rangle^n$. This is achieved by solving a SAT formula $\mathcal{J}_K=\mathcal{J}\langle I,F,E \rangle (P_{a-k},\mat{T})$ that satisfies Jury's criteria for $P_{a-k}=z^n+\sum_{i=1}^n (a_i-k_i) z^{n-i} : k_i \in \mat{K} \wedge \mat{K}=\tilde{\mat{K}}\mat{T}$.
\item Select the observer error characteristic polynomial $\tilde{\mat{L}} \in \mathbb{R}\langle I,F,E\rangle^n$. 
This is achieved by solving a SAT formula $\mathcal{J}_L=\mathcal{J}\langle I,F,E \rangle (P_p,\mat{W}_o^{-1}\mat{W}_{of})$ 
that satisfies Jury's criteria for 
$P_p=z^n+\sum_{i=1}^n p_i z^{n-i} : p_i \in \vec{P}= (\mat{W}_o^{-1}\mat{W}_{of})^-1 \tilde{\mat{L}} 
+ [\begin{array}{cccc}a_1&a_2&\hdots&a_n\end{array}]^T$.
\item If there is no candidate solution we return UNSAT and exit the loop.
\end{enumerate}
\item Once we have a candidate solution, we perform a safety verification using abstract acceleration.
\begin{enumerate}
\item Calculate the abstract matrices $\mathcal{A}_t, \mathcal{B}_{tp},\mathcal{B}_{tn}$ for the candidate closed loop solution.
\item Evaluate $\hat{X}^\# \models \Psi$. If the verification holds we have a solution, and exit the loop.
\end{enumerate} 
\item If the candidate solution is not valid we refine the abstraction used by the synthesiser
\begin{enumerate}
\item Find the constraints that invalidate the property. These are the rows of $\mat{E} \in \Psi$ that intersect/segment $\hat{X}^\#$. Transforming these rows into the eigenspace of $\mat{A}_t$, and applying backward acceleration to obtain a set of counterexamples for the eigenvalues, which we define as $\Psi_{\lambda}$ .
\item Transform the soft constraints on the eigenvalues defined in the counterexample into soft constraints on the coefficients of the characteristic polynomial. 
$z^n+\sum_{i=1}^n[a_i-k_i]z^{n-i}=\prod_{i=1}^n (z-[\lambda_i]) : |[\lambda_i]|<1 \wedge [\lambda_i] \models \Psi_{\lambda}$.
\item Pass the refined abstraction with the new contraints to the synthesiser and repeat the loop again.

\end{enumerate} 
\end{enumerate}

\begin{figure}
\centering
\hspace*{-1cm}
\begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, ampersand replacement=\&,]
  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center},row sep=.2cm, column sep=.2cm] {
   \coordinate (aux1);
   \& \coordinate (aux2);
   \& ;\\
   \&\node[fill=yellow!20,align=center] (pre) {{\sc pre-}\\{\sc processing}};
   \& \node[fill=yellow!20,align=center] (abstract) {\sc abstraction};
        \node[draw=none] (FIK) at ([xshift=2.5cm,yshift=.3cm]abstract)  {\sc $\Psi (\mat{K})$};
   \& \node[fill=yellow!20,align=center] (counter) {\sc counterexample};\\
   \& 
   \& \node[fill=yellow!20,align=center, minimum width=3cm] (synth) {\sc synthesiser};
       \node[draw=none] (KL) at ([xshift=2.5cm,yshift=-.25cm]synth)  {\sc $\mat{K},\mat{L}$};
       \node[draw=none] (NKL) at ([xshift=2cm,yshift=1.2cm]synth)  {\sc $\neg \mat{K}, \mat{L}, \mat{A}$};
   \& \node[fill=yellow!20,align=center, minimum width=4cm] (verify) {\sc verifier ($\Psi$)};
       \node[draw=none] (VUNS) at ([xshift=.6cm,yshift=1cm]verify)  {\sc UNSAT};
       \node[draw=none] (SAT) at ([xshift=2.5cm,yshift=-.3cm]verify)  {\sc SAT};
   \& \node[ellipse, fill=yellow!20] (done) {{\sc Done}};\\
   \&
   \& complexnode/.pic={
     \coordinate (auxsynth);
     \node[draw,rectangle,align=center] (KSAT) at ([xshift=-.8cm]auxsynth.center) {\sc SAT\\Solver};
     \node[draw,rectangle,align=center] (LSAT) at ([xshift=.8cm]auxsynth.center) {\sc SAT\\Solver};
     \node[draw=none] (SATK) at ([xshift=-.5cm,yshift=1.5cm]KSAT)  {$\mathcal{J}_K$};
     \node[draw=none] (K) at ([xshift=.4cm,yshift=1.5cm]KSAT)  {$\tilde{\mat{K}}$};
     \node[draw=none] (SATL) at ([xshift=-.5cm,yshift=1.5cm]LSAT)  {$\mathcal{J}_L$};
     \node[draw=none] (L) at ([xshift=.4cm,yshift=1.5cm]LSAT)  {$\tilde{\mat{L}}$};
    }
   \& complexnode/.pic={
     \coordinate (AA);
     \node[draw,rectangle,align=center] (AAI) at ([xshift=-1.5cm]AA.center) {\sc AI\\Model};
     \node[draw,rectangle,align=center] (AAV) at ([xshift=1.5cm]AA.center) {\sc AA\\Solver};
     \node[draw=none] (AB) at ([xshift=.4cm,yshift=1.5cm]AAI.center)  {\sc $\mat{A}_t$ $\mat{B}_{tp}$ $\mat{B}_{tn}$};     
     \node[draw=none] (MA) at ([yshift=.2cm]AA.center)  {$\mathcal{A}_t$};     
     \node[draw=none] (MB) at ([yshift=-.3cm]AA.center)  {$\mathcal{B}_{tp},\mathcal{B}_{tn}$}; 
     \node[draw=none] (AB) at ([xshift=0.3cm,yshift=1.5cm]AAV.center)  {\sl $\hat{X}^\#$};  
    }\\
   \&
   \& complexnode/.pic={
     \coordinate (UNSAT);
     \coordinate (KUNSAT)  at ([xshift=-.8cm]UNSAT.center);
     \coordinate (LUNSAT) at ([xshift=.8cm]UNSAT.center);
     \node[draw=none] (UNSATB) at ([xshift=1.6cm,yshift=.3cm]UNSAT)  {\sc UNSAT};
  }
  \& \& \coordinate (FAIL);\\
  };
  \path
    (pre.east) edge (abstract.west)
    (abstract.south) edge (synth.north)
    (synth.east) edge (verify.west)
    (verify.east) edge (done.west);
  \path
    ([xshift=+.6cm]KSAT.north) edge ([xshift=-2.1cm]synth.south) 
    ([xshift=-3.3cm]synth.south) edge ([xshift=-.6cm]KSAT.north)
    ([xshift=.6cm]LSAT.north) edge ([xshift=3.3cm]synth.south)
    ([xshift=+2.1cm]synth.south) edge ([xshift=-.6cm]LSAT.north)
    ([xshift=-4.9cm]verify.south) edge (AAI.north)
    (AAI.east) edge (AAV.west)
    (AAV.north) edge ([xshift=4.9cm]verify.south)
    (verify.north) edge (counter.south)
    (counter.west) edge (abstract.east)
    (counter.south west) edge (synth.north east)
    (FAIL.north) edge (done.south);
   \path[-] 
     (KSAT.south) edge (KUNSAT.north)
     (LSAT.south) edge (LUNSAT.north)
     (KUNSAT.east) edge (LUNSAT.west)
     (LUNSAT.east) edge (FAIL.west);
\end{tikzpicture}
\caption{The CEGIS loop with CEGAR\label{fig:CEGARIS}.}
\end{figure}

@@

\paragraph{Discretization Effects}

The classical approach to control synthesis has often disregarded
quantization as a minimal effect, thus proving unsound in the case of digital control systems.
Modern techniques focus on different elements of the discretization effect, including delayed
response~\cite{Duggirala2015}, and Finite Word Length (FWL) semantics with
the goal to either verify (e.g.~\cite{daes20161}) or optimize
(e.g.~\cite{oudjida2014design}) given implementations.
